---
title: Traffic Data Analysis
output: html_document
---

## Loading necessary libraries

```{r}
library(dplyr)
library(ggplot2)
```
## Loading the data:

```{r}
data <- read.csv('all-semi-unique.csv')
```

## Getting basic information about the data:

```{r}
glimpse(data)
View(data)
names(data)
dim(data)
```

Hence, from applying the above functions, we have a genral idea about the data: how the different records look like, the number of records (rows), the number of columns and their corresponding names.

Also let's see the number of NAs for each column:

```{r}
sapply(data, function(d) sum(is.na(d)))
```

## Cleaning the data:

Let's see the unique values in each column:

```{r}
sapply(data, function(d) length(unique(d)))
```

We can see now that all the "ad" columns (columns with name starting with "ad.") have only 1 unique value. Also the columns "rd.cl" and "rd.rp.type" have only 1 unique value.

Those are useless columns for our analysis. So let's remove them:

```{r}
data <- select(data, -starts_with("ad."))
data <- select(data, -c(rd.cl, rd.rp.type))
```

Now let's see the updated number of columns:

```{r}
ncol(data)
```

Let's check for duplicate rows:

```{r}
nrow(data) - nrow(unique(data))
```

We can see that there are no duplicate rows present in the dataset.